Can Zero-shot LLMs save additional work in machine learning prioritised screening for systematic reviews?

### 2024-11-01 LLMs/pres/2024_10_30_LLM-screening.tex
manuscript updates

\documentclass[9pt,aspectratio=169]{beamer}
\mode<presentation>
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{listings}

@@

\title[LLM screening and coding]{Screening and coding with LLMs}

@@

\begin{frame}{Outlook}
	\begin{itemize}
		\item<1-> Results from the implementation of \cite{wang_zero-shot_2024}
		\item<2-> Additional results from Santiago's thesis project on coding
	\end{itemize}
\end{frame}

@@

\section{Screening}

\begin{frame}{We can ask LLMs whether a study should be included in a systematic review}
	\begin{columns}
		\begin{column}{0.5\linewidth}
			\begin{figure}
				\footnotesize
				%\includegraphics[width=\linewidth]{images/example.txt}
				\input{images/example.txt}
			\end{figure}
		\end{column}
		\begin{column}{0.5\linewidth}
			\only<2->{\begin{figure}
					\includegraphics[width=\columnwidth]{images/example_probs.pdf}
			\end{figure}}
		\end{column}
	\end{columns}
\end{frame}

@@

\begin{frame}{Results are mixed, because the probabilities are poorly calibrated, and their attempts to re-invent stopping criteria do not makes sense}

	\begin{figure}
		\includegraphics[width=0.8\columnwidth]{images/tab3-1}
	\end{figure}
\end{frame}

@@

\begin{frame}[fragile]{Implementation is relatively straightforward}
	\small
\begin{lstlisting}[language=Python]
prompt = Template('''<s>[INST] <<SYS>>
You are a systematic review helper tasked with finding out whether a study is relevant to the review $t

@@

%\section{Results}

\begin{frame}{Result \#1: It works!}
	\begin{columns}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item<1-> The method produces a ranking that identifies all relevant documents before all documents have been screened
				\item<2-> But it doesn't perform as well as simple tfidf + SVM
				\item<3-> To compare the quality of rankings, we can use ROC AUC scores
			\end{itemize}
		\end{column}
		\begin{column}{0.6\linewidth}
			\only<1>{\includegraphics[width=\linewidth]{../../figures/LLM_Brouwer_simple.pdf}}
			\only<2->{\includegraphics[width=\linewidth]{../../figures/Brouwer_2019.pdf}}
			%\only<3>{}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{This method simply produces a ranking of documents, but is this ranking better than what we can produce with traditional active learning approaches?}

	\begin{columns}
		\begin{column}{0.618\linewidth}
			\begin{figure}

				\only<1>{\includegraphics[width=\columnwidth]{../../figures/Appenzeller-Herzog_2019_p_distribution.pdf}}
				\only<2>{\includegraphics[width=\columnwidth]{../../figures/Appenzeller-Herzog_2019.pdf}}
				\only<3>{\includegraphics[width=\columnwidth]{../../figures/Bos_2018.pdf}}
				\only<4>{\includegraphics[width=\columnwidth]{../../figures/Brouwer_2019.pdf}}
				\only<5>{\includegraphics[width=\columnwidth]{../../figures/Chou_2004.pdf}}
				\only<6>{\includegraphics[width=\columnwidth]{../../figures/Donners_2021.pdf}}
				\only<7>{\includegraphics[width=\columnwidth]{../../figures/Jeyaraman_2020.pdf}}
			\end{figure}
		\end{column}
		\begin{column}{0.382\linewidth}
			\begin{itemize}
				\item<1-> We generated LLM inclusion scores for each document in a set of reviews in the Synergy dataset
				\item<2-> We compare what happens to screening in descending order of this ranking to 100 active learning runs with a default configuration of SVM with Tfidf
				\item<3-> Initial results are not very promising
				\item<5->But LLMs can do well where traditional methods struggle
				\item<7->Sometimes they are no use at all
			\end{itemize}
		\end{column}
	\end{columns}

@@

\begin{frame}{All results}
	\begin{columns}
	\begin{column}{0.618\linewidth}
		\begin{figure}

			\only<1->{\includegraphics[width=\columnwidth]{../../figures/llm_svm_roc.pdf}}

		\end{figure}
	\end{column}
	\begin{column}{0.382\linewidth}
		\begin{itemize}
			\item<1-> LLMs and basic active learning pipelines seem to have different weakneses
			\item<2-> Combining both could improve general performance
			\item<3-> LLMs seem most useful for smaller datasets (where active learning has little time to learn)
		\end{itemize}
	\end{column}
\end{columns}
\end{frame}

@@

\begin{itemize}
	\item<1-> Prompting strategies (inclusion criteria)
	\item<2-> Bigger/different models
	\item<3-> Combining LLMs with traditional approaches
	\item<4-> Updating prompts based on user feedback
\end{itemize}

@@

\begin{frame}{Conclusions}
	\begin{columns}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item<1-> LLMs are neither a quick fix or a silver bullet
				\item<2-> Evaluation is vital
				\item<3-> We can't forget the need for appropriate stopping criteria \cite{callaghan_statistical_2020}
				%\item<4->\url{mcallaghan.github.io/buscar-app}
			\end{itemize}
		\end{column}
		\begin{column}{0.6\linewidth}
			\begin{figure}
				\only<3>{\includegraphics[width=\columnwidth]{../../figures/stopping.pdf}}
			\end{figure}
		\end{column}
	\end{columns}

@@

\section{Coding}

\begin{frame}{Using LLMs for coding}
		\begin{columns}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item<1-> Santiago's thesis showed us that LLMs can achieve comparable performance with BERT-type models
				\item<2->They need no training data to achieve this performance
				\item<3->But we do need annotated data for ``prompt engineering'' ($\approx$ training?), and for evaluation
			\end{itemize}
		\end{column}
		\begin{column}{0.6\linewidth}
			\begin{figure}
				\only<1->{\includegraphics[width=\columnwidth]{../../figures/LLM-coding.pdf}}
			\end{figure}
		\end{column}
	\end{columns}

\begin{frame}{Using LLMs for coding}
	\begin{columns}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item<1-> Results were aggregated Macro F1 scores for 30 labels
				\item<2-> Aggregate results hide much variation
				\item<3-> This varies across the different coding categories
				\item<4-> And there seems to be a clear correlation with number of labels
			\end{itemize}
		\end{column}
		\begin{column}{0.6\linewidth}
			\begin{figure}
				\only<2>{\includegraphics[width=\columnwidth]{../../../LLM-eval/gpt/figures/BERT-vs-LLMs.pdf}}
				\only<3>{\includegraphics[width=\columnwidth]{../../../LLM-eval/gpt/figures/BERT-vs-LLMs-categories.pdf}}
				\only<4>{\includegraphics[width=\columnwidth]{../../../LLM-eval/gpt/figures/BERT-vs-LLMs-n.pdf}}
			\end{figure}
		\end{column}
	\end{columns}

	\begin{frame}{Conclusions}
\end{frame}
%	
%\begin{frame}
%	Thanks!
%	
%	\bigskip
%	
%	\hrule
%	
%	\bigskip
%	
%	\url{callaghan@mcc-berlin.net}
%	 
%	\medskip
%	
%	Twitter: @MaxCallaghan5
%	
%	\bigskip
%	
%	\begin{figure}
%	\includegraphics[width=0.1\columnwidth]{images/wellcome-logo-black.eps}
%	\end{figure}
%\end{frame}
%

>>>>

\documentclass[9pt,aspectratio=169]{beamer}
\mode<presentation>
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{listings}
\usepackage{booktabs}

@@

\title[LLM screening]{Can Zero-shot LLMs save additional work in machine learning prioritised screening for systematic reviews?}

@@

@@

\section{Introduction}

\begin{frame}{Can LLMs help us to save more work?}

\begin{columns}
	\begin{column}{0.5\linewidth}
		\begin{itemize}
			\item<1->Prioritised screening has saved us a lot of work
			\item<2->But we still spend substantial amounts of time screening documents for systematic reviews
			\item<3->LLMs have been proposed as a solution, with
			\begin{itemize}
				\item<4-> \cite{wang_zero-shot_2024} showing impressive results for the zero-shot setting
				\item<5-> and  \cite{xia_llmscreen_2024} erroneously claiming that LLMs fix all of the problems faced until now by ml+screening
			\end{itemize} 
			\item<6-> There are a lot of bad papers, which will continue to proliferate, each tweaking some parameters to achieve a high performance that cannot be generalised
		\end{itemize}
	\end{column}

	\begin{column}{0.5\linewidth}
		\only<4>{
		\begin{figure}
			\includegraphics[width=\columnwidth]{images/tab3-1}
		\end{figure}
		}	
		\only<5->{
					\begin{figure}
				\includegraphics[width=\columnwidth]{images/xia}
			\end{figure}
	}
	\end{column}
\end{columns}

\begin{frame}{Realistic and useful evaluations of LLMs for screening}

	Despite the hype, LLMs might still be helpful, but we need to know

	\begin{itemize}
		\item<2->How do LLMs compare against commonly used (and much cheaper!) baselines?
		\item<3->How could they safely be used in the real world (Hint: with stopping criteria and prioritised screening)
	\end{itemize}
\end{frame}

\section{Prioritisation with LLMs}

@@

@@

\begin{frame}[fragile]{Implementation is relatively straightforward}

	\small
\begin{lstlisting}[language=Python]
prompt = Template('''<s>[INST] <<SYS>>
You are a systematic review helper tasked with finding out whether a study is relevant to the review $t

@@

\begin{frame}[fragile]{Implementation is relatively straightforward}

	\begin{columns}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\footnotesize
			%\includegraphics[width=\linewidth]{images/example.txt}
			\input{images/example.txt}
		\end{figure}
	\end{column}
	\begin{column}{0.5\linewidth}
		\only<2->{\begin{figure}
				\includegraphics[width=\columnwidth]{images/example_probs.pdf}
		\end{figure}}
	\end{column}
\end{columns}

@@

\section{Comparison}

\begin{frame}{Comparing priorisation approaches}

\begin{itemize}
	\item<1-> We evaluate using the Synergy dataset \cite{de_bruin_synergy_2023} of 26 systematic reviews (Tim, Diana, Sergio, Lena, and James, will make a much bigger one!)
	\item<2-> For each review, we generate 100 different initial random samples of 10\%, and do normal prioritised screening in batches of 10\% with an SVM 100 times
	\item<3-> We also make a prediction for each document with 4 Llama models of different sizes and ages
	\item<4-> We turn these predictions into a prioritised list of documents to be screened
\end{itemize}

\section{Results}

\begin{frame}{Result \#1: It works!}
	\begin{columns}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item<1-> The method produces a ranking that identifies all relevant documents before all documents have been screened
				\item<2-> The first model we tried was much worse than our SVM, but newer and bigger models were much more promising!
				\item<3-> But performance varied across datasets
				\item<4-> ROC AUC scores offer one way to compare rankings
			\end{itemize}
		\end{column}
		\begin{column}{0.6\linewidth}
			\only<1>{\includegraphics[width=\linewidth]{../../figures/LLM_Brouwer_simple.pdf}}
			\only<2>{\includegraphics[width=\linewidth]{../../figures/Brouwer_2019.pdf}}
				\only<3>{\includegraphics[width=\columnwidth]{../../figures/Bos_2018.pdf}}
				\only<4>{\includegraphics[width=\columnwidth]{../../figures/Appenzeller-Herzog_2019.pdf}}
			\only<5>{\includegraphics[width=\columnwidth]{../../figures/Chou_2004.pdf}}
			\only<6>{\includegraphics[width=\columnwidth]{../../figures/Donners_2021.pdf}}
			\only<7>{\includegraphics[width=\columnwidth]{../../figures/Jeyaraman_2020.pdf}}
			\only<8>{\includegraphics[width=\columnwidth]{../../figures/macro_comparison.pdf}}
			%\only<3>{}
		\end{column}
	\end{columns}
\end{frame}

%\begin{frame}{All results}
%	\begin{columns}
%	\begin{column}{0.618\linewidth}
%		\begin{figure}
%			
%			\only<1->{\includegraphics[width=\columnwidth]{../../figures/llm_svm_roc.pdf}}
%
%		\end{figure}
%	\end{column}
%	\begin{column}{0.382\linewidth}
%		\begin{itemize}
%			\item<1-> LLMs and basic active learning pipelines seem to have different weakneses
%			\item<2-> Combining both could improve general performance
%			\item<3-> LLMs seem most useful for smaller datasets (where active learning has little time to learn)
%		\end{itemize}
%	\end{column}
%\end{columns}
%\end{frame}

\begin{frame}{Stopping criteria}
	\begin{columns}
		\begin{column}{0.32\linewidth}
			\begin{itemize}
				\item<1->Next, we calculate scores for our stopping criteria with a range of different recall targets
				\item<2->Stopping when $p<\alpha$ gives a recall > target in the $\alpha^{th}$ percentile, in 97/100 settings
				\item<3->In most settings, total work savings were greater for SVMs 
			\end{itemize}
		\end{column}
		\begin{column}{0.68\linewidth}
			\only<2>{
			\begin{table}
				\scriptsize
				\input{../../tables/recall.tex}

			\end{table}
		}
			\only<3>{
	\begin{table}
		\scriptsize
		\input{../../tables/wsst.tex}
						\caption{absolute work savings across datasets (N records=169,288)}
	\end{table}
}
		\end{column}
	\end{columns}
\end{frame}

@@

\begin{itemize}
	\item<1-> Prompting strategies (inclusion criteria)
	\item<2-> Combining LLMs with traditional approaches
	\item<3-> Updating prompts based on user feedback
	\item<4-> Better baselines (finetuning BERT)
\end{itemize}

\only<5->{Note that although the LLM implementation is simplistic, so is the comparator}

@@

\begin{frame}{Conclusions}
	\begin{columns}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item<1-> LLMs are neither a quick fix or a silver bullet
				\item<2-> They do not have magic properties that fix problems of traditional ML, they are simply likely to make them worse
				\item<3-> We need to be on our guard!
				\item<4-> Do not forget stopping criteria!
				\item<5-> But they have some advantages, such as the fact that we only need to calculate scores once!
				%\item<4->\url{mcallaghan.github.io/buscar-app}
			\end{itemize}
		\end{column}
		\begin{column}{0.6\linewidth}
			\begin{figure}
				\only<4>{\includegraphics[width=\columnwidth]{../../figures/stopping.pdf}}
			\end{figure}
		\end{column}
	\end{columns}

@@

%	
%\begin{frame}
%	Thanks!
%	
%	\bigskip
%	
%	\hrule
%	
%	\bigskip
%	
%	\url{callaghan@mcc-berlin.net}
%	 
%	\medskip
%	
%	Twitter: @MaxCallaghan5
%	
%	\bigskip
%	
%	\begin{figure}
%	\includegraphics[width=0.1\columnwidth]{images/wellcome-logo-black.eps}
%	\end{figure}
%\end{frame}
%
